Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields

### 2017-01-04 Master Thesis/synopsis.tex
Show that the big-ass set is closed

\begin{definition}
Let $K$ be finitely generated. A \textit{model} of $K$ is a regular, proper scheme $\mathfrak{X}$ over $\mathbb{Z}$ with function field $K$. A \textit{morphism of models} is a morphism of schemes inducing the identity on $K$. We denote the category of models of $K$ by $\mathfrak{M}(K)$; it is filtered.

@@

\begin{lemma}\label{2.4-pre}
Let $K$ be finitely generated, $\ell\neq 2$ be a prime and $f\in\mathrm{H}^3(K, \mathbb{Z}/\ell\mathbb{Z})$. Then the support of $f$ is closed in $|\operatorname{Spv}(K)|$:
\[ \operatorname{supp}(f) = \{w\in |\operatorname{Spv}(K)|\mid \operatorname{res}_{K_w/K}(f)\neq 0 \} \]
\end{lemma}

\begin{lemma}\label{2.4}
Let $\ell\neq 2$ be a prime. The following map, induced by the restrictions, is injective:
\[ \mathrm{H}^3(K,\mathbb{Z}/\ell\mathbb{Z})\to\prod_{w\in |\operatorname{Spv}(K)| }\mathrm{H}^3(K_w,\mathbb{Z}/\ell\mathbb{Z}) \]
\end{lemma}

@@

First, let's consider the case where $K$ is finitely generated. Let $\mathfrak{X}$ be a model of $K$. $\mathfrak{X}_i$ denotes the set of points $x\in\mathfrak{X}$ whose closure $\overline{\{x\}}$ is $i$-dimensional. For $x\in\mathfrak{X}_1$ the local ring $\mathcal{O}_{\mathfrak{X}, x}$ is a discrete valuation ring and defines a valuation $v_x\in\operatorname{Spv}_\text{d}^{(1)}(K)$. By [K1] Theorem 0.8, the following map
\[ \mathrm{H}^3(K)\longrightarrow \prod_{x\in\mathfrak{X}_1}\mathrm{H}^3(K_{v_x}) \]
is injective and its image is contained in $\bigoplus \mathrm{H}^3(K_{v_x})$. By [K1] Lemma 1.4, for a discrete valuation $v$ of $K$ the map
\[ \mathrm{H}^3(K_v) \longrightarrow \prod_w\mathrm{H}^3(K_w) \]
where $w$ ranges over all rank-$2$ valuations with $v(w) = v$, can be identified with the following map:
\[ {}_\ell\operatorname{Br}(k) \longrightarrow \prod_{\tilde{w}\in\operatorname{Spv}^{(1)}(k)} {}_\ell\operatorname{Br}(k_{\tilde{w}}) \]
where $k = k(v)$. This map is injective by the classical Hasse principle and its image is contained in $\bigoplus {}_\ell\operatorname{Br}(k_{\tilde{w}})$. Combining these maps yields the injection:
\[ \mathrm{H}^3(K)\longrightarrow \bigoplus_{x\in\mathfrak{X}_1}\  \bigoplus_{\substack{w\in\operatorname{Spv}_\text{rd}^{(2)}(K) \\ v(w) = v_x}} \mathrm{H}^3(K_w) \]
The direct sum ranges over a subset of $|\operatorname{Spv}(K)|$, thus this shows the assertion. Now, we drop the assumption that $K$ is finitely generated.

@@

Let  $w\in |\operatorname{Spv}(K)|$. Since $\mathrm{H}^3(K_w) = \varinjlim_I \mathrm{H}^3((K_i)_w)$ there exists an $i = i(w)\in I$ such that $\operatorname{res}_{(K_i)_w/K}(g) = 0$. By Lemma \ref{2.4-pre} there is an open neighborhood $U_{i(w)}\subset |\operatorname{Spv}(K_i)|$ of $w$ such that $\operatorname{res}_{(K_i)_v/K}(g)=0$ for all $v\in U_{i(w)}$. For $K_i\subset K_j\subset K$ let
\[ \varphi_i: |\operatorname{Spv}(K)| \to |\operatorname{Spv}(K_i)| \]
\[ \varphi_{ij}: |\operatorname{Spv}(K_j)|\to |\operatorname{Spv}(K_i)| \]
be the restriction maps. They are continuous. Since $|\operatorname{Spv}(K)|$ is quasicompact we can find a finite set $S\subset |\operatorname{Spv}(K)|$ such that:
\[ \bigcup_{w\in S}\varphi^{-1}_{i(w)}(U_{i(w)}) = |\operatorname{Spv}(K)| \]
Let $j\in I$ such that $K_j$ is the composite of the fields $K_{i(w)},\ w\in S$. We also have:
\[ \bigcup_{w\in S}\varphi^{-1}_{i(w)j}(U_{i(w)}) = |\operatorname{Spv}(K_j)| \]
Let $v\in |\operatorname{Spv}(K_j)|$ and let $i = i(w),\ w\in S$ such that $v\in\varphi_{ij}^{-1}(U_{i(w)})$. We have:
\[ \operatorname{res}_{(K_j)_v/K_j}(\operatorname{res}_{K_j/K_0}(g)) = \operatorname{res}_{(K_j)_v/(K_i)_v}(\operatorname{res}_{(K_i)_v/K_0}(g)) = 0 \]
Since $K_j$ is finitely generated and we proved the statement for finitely generated fields, we obtain $\operatorname{res}_{K_j/K_0}(g) = 0$, thus $f = \operatorname{res}_{K/K_0}(g) = 0$.
\end{proof}

>>>>

\begin{definition}
Let $K$ be finitely generated. A \textit{model} of $K$ is a regular, proper scheme $\mathfrak{X}$ over $\mathbb{Z}$ with function field $K$. $\mathfrak{X}_i$ denotes the set of points $x\in\mathfrak{X}$ whose closure $\overline{\{x\}}$ is $i$-dimensional.

A \textit{morphism of models} is a morphism of schemes inducing the identity on $K$. We denote the category of models of $K$ by $\mathfrak{M}(K)$; it is filtered.

@@

\begin{lemma}\label{2.4-finitelygenerated}
Let $K$ be finitely generated and $\ell\neq 2$ be a prime. The following map, induced by the restrictions, is injective:
\[ \mathrm{H}^3(K,\mathbb{Z}/\ell\mathbb{Z})\to\prod_{w\in |\operatorname{Spv}(K)| }\mathrm{H}^3(K_w,\mathbb{Z}/\ell\mathbb{Z}) \]
\end{lemma}

@@

Let $\mathfrak{X}$ be a model of $K$. For $x\in\mathfrak{X}_1$ the local ring $\mathcal{O}_{\mathfrak{X}, x}$ is a discrete valuation ring and defines a valuation $v_x\in\operatorname{Spv}_\text{d}^{(1)}(K)$. By [K1] Theorem 0.8, the map
\[ \mathrm{H}^3(K)\longrightarrow \prod_{x\in\mathfrak{X}_1}\mathrm{H}^3(K_{v_x}) \]
is injective and its image is contained in $\bigoplus \mathrm{H}^3(K_{v_x})$. By [K1] Lemma 1.4, for a discrete valuation $v$ of $K$ the map
\[ \mathrm{H}^3(K_v) \longrightarrow \prod_w\mathrm{H}^3(K_w) \]
where $w$ ranges over all rank-$2$ valuations with $v(w) = v$, can be identified with the following map:
\[ {}_\ell\operatorname{Br}(k) \longrightarrow \prod_{\tilde{w}\in\operatorname{Spv}^{(1)}(k)} {}_\ell\operatorname{Br}(k_{\tilde{w}}) \]
where $k = k(v)$. This map is injective by the classical Hasse principle and its image is contained in $\bigoplus {}_\ell\operatorname{Br}(k_{\tilde{w}})$. Combining these maps yields the injection:
\[\Phi: \mathrm{H}^3(K)\longrightarrow \bigoplus_{x\in\mathfrak{X}_1}\  \bigoplus_{\substack{w\in\operatorname{Spv}_\text{rd}^{(2)}(K) \\ v(w) = v_x}} \mathrm{H}^3(K_w) \]
The direct sum ranges over a subset of $|\operatorname{Spv}(K)|$, thus this concludes the proof. 
\end{proof}

\begin{lemma}\label{2.4-closedset}
Let $K$ be finitely generated, $\ell\neq 2$ be a prime and $f\in\mathrm{H}^3(K)$. Let $\mu_\ell\subset K$ and $K$ not formally real. Then the support of $f$ is closed in $|\operatorname{Spv}(K)|$:
\[ \operatorname{supp}(f) = \{w\in |\operatorname{Spv}(K)|\mid \operatorname{res}_{K_w/K}(f)\neq 0 \} \]
where $\operatorname{res}_{K_w/K}: \mathrm{H}^3(K)\to \mathrm{H}^3(K_w)$.
\end{lemma}

\begin{proof}
We adopt the notation from the previous proof. First, we shall prove that the following map is injective for every model $\mathfrak{X}$ of $K$:
\[ f: \mathrm{H}^3(K) \to \prod_{y\in\mathfrak{X}_0}\mathrm{H}^3(K_y) \]
For $x\in\mathfrak{X}$ let $\mathcal{O}_x$ be a henselization of $\mathcal{O}_{\mathfrak{X}, x}$ and $K_x$ its quotient field. For $y\in\mathfrak{X}_0$ set $P_y=\operatorname{Spec}(\mathcal{O}_y)_1$. For $z\in P_y$ let $\mathcal{O}_z$ be a henselization of $\mathcal{O}_y$ at $z$ and $K_z$ its quotient field. For $y\in\mathfrak{X}_0$ and $x\in\mathfrak{X}_1$ let $P_{x,y}$ be the subset of $P_y$ consisting of all $z\in P_y$ which are mapped to $x$ under $\operatorname{Spec}(\mathcal{O}_y)\to\mathfrak{X}$. Consider the following commutative diagram:
\[ \begin{tikzcd}
\mathrm{H}^3(K)\ar[r, "\varphi"]\ar[d, "f"'] & \displaystyle\prod_{x\in\mathfrak{X}_1}\mathrm{H}^3(K_x)\ar[d, "\psi"]\\
\displaystyle\prod_{y\in\mathfrak{X}_0}\mathrm{H}^3(K_y) \ar[r, "g"']& \displaystyle\prod_{\substack{x\in\mathfrak{X}_1 \\ y\in\mathfrak{X}_0}}\prod_{z\in P_{x,y}}\mathrm{H}^3(K_z)
\end{tikzcd} \]
According to [Sa], Theorem (2-15) $g$ is injective. Now, $\psi\circ\varphi$ can be identified with the injective map $\Phi$ from the proof of the previous Lemma. Thus its image is contained in $\bigoplus\mathrm{H}^3(K_w)$. Therefore, $f$ is injective as well and its image is contained in $\bigoplus\mathrm{H}^3(K_y)$. Now, one easily checks that
\[ \operatorname{supp}(f) = \bigcap_{\mathfrak{M}_K} \pi_\mathfrak{X}^{-1}(\{ y\in\mathfrak{X}_0\mid \operatorname{res}_{K_y/K}(f)\neq 0 \}) \]
is closed.
\end{proof}

\begin{lemma}\label{2.4}
Let $K$ be not necessarily finitely generated and $\ell\neq 2$ be a prime. The following map, induced by the restrictions, is injective:
\[ \mathrm{H}^3(K,\mathbb{Z}/\ell\mathbb{Z})\to\prod_{w\in |\operatorname{Spv}(K)| }\mathrm{H}^3(K_w,\mathbb{Z}/\ell\mathbb{Z}) \]
\end{lemma}

\begin{proof}
Just like in Lemma \ref{2.4-finitelygenerated} we can assume that $\mu_\ell\subset K$ and $K$ is not formally real and replace $\mathbb{Z}/\ell\mathbb{Z}$ by $\mathbb{Z}/\ell\mathbb{Z}(2)$. We adopt the notation from the previous proof.

@@

Let  $w\in |\operatorname{Spv}(K)|$. Since $\mathrm{H}^3(K_w) = \varinjlim_I \mathrm{H}^3((K_i)_w)$ there exists an $i = i(w)\in I$ such that $\operatorname{res}_{(K_i)_w/K}(g) = 0$. By Lemma \ref{2.4-closedset} there is an open neighborhood $U_{i(w)}\subset |\operatorname{Spv}(K_i)|$ of $w$ such that $\operatorname{res}_{(K_i)_v/K}(g)=0$ for all $v\in U_{i(w)}$. For $K_i\subset K_j\subset K$ let
\[ \varphi_i: |\operatorname{Spv}(K)| \to |\operatorname{Spv}(K_i)| \]
\[ \varphi_{ij}: |\operatorname{Spv}(K_j)|\to |\operatorname{Spv}(K_i)| \]
be the restriction maps. They are continuous. Since $|\operatorname{Spv}(K)|$ is quasicompact we can find a finite set $S\subset |\operatorname{Spv}(K)|$ such that:
\[ \bigcup_{w\in S}\varphi^{-1}_{i(w)}(U_{i(w)}) = |\operatorname{Spv}(K)| \]
Let $j\in I$ such that $K_j$ is the composite of the fields $K_{i(w)},\ w\in S$. We also have:
\[ \bigcup_{w\in S}\varphi^{-1}_{i(w)j}(U_{i(w)}) = |\operatorname{Spv}(K_j)| \]
Let $v\in |\operatorname{Spv}(K_j)|$ and let $i = i(w),\ w\in S$ such that $v\in\varphi_{ij}^{-1}(U_{i(w)})$. We have:
\[ \operatorname{res}_{(K_j)_v/K_j}(\operatorname{res}_{K_j/K_0}(g)) = \operatorname{res}_{(K_j)_v/(K_i)_v}(\operatorname{res}_{(K_i)_v/K_0}(g)) = 0 \]
Since $K_j$ is finitely generated we obtain $\operatorname{res}_{K_j/K_0}(g) = 0$ with Lemma \ref{2.4-finitelygenerated}, thus $f = \operatorname{res}_{K/K_0}(g) = 0$.
\end{proof}
